
```{r methods-data, include=FALSE}
fp <- here("analysis", "data", "derived_data", "mids_rf.rds")
mids_rf <- readr::read_rds(fp)
```

### Animals
Female Sprague-Dawley rats aged 10-12 weeks (220-250 g) at the start of the experiment were used in this study. Rats were grouped-housed in standard polycarbonate cages on a 12 h light-dark cycle (lights on 08:00-20:00), standard temperature (22 ± 2`r knitr::asis_output("\u00B0C")`) and humidity (40-55%) with food and water available _ad libitum_. Experiments were approved by the Ethical Committee for Animal Experiments of the Estonian Ministry of Rural Affairs (permission 25.06.2018 nr 127) in accordance with Directive 2010/63/EU of the European Union.

### Drugs
Racemic ketamine (brand name, company – Marianna ) diluted in sterile saline was administered intraperitoneally at a dose of 30 mg/kg and an injection volume of 1 ml/kg  (check injection volume – Marianna) once per day for 10 consecutive days. Control animals received saline at 1 ml/kg.

### Tissue Collection
Brains were collected 24h after the last injection and flash frozen in isopentane on dry ice followed by storage at -80`r knitr::asis_output("\u00B0C")`. Sections of 40 µm were sliced on a Leica CM1520 cryostat onto VWR Superfrost Plus slides and stored at -80`r knitr::asis_output("\u00B0C")` until use. 

### Cytochrome C oxidase staining and image analysis
Cytochrome C oxidase staining and image analysis were performed as described in Kanarik and Harro [-@kanarikSociabilityTraitRegional2018], based on a modified protocol by Gonzalez-Lima and Cada [-@gonzalez-limaQuantitativeHistochemistryCytochrome1998]. Enzyme activity levels were derived from optical density measurements of a histochemical reaction product within each brain region. The optical density values were converted to enzyme activity levels by using external standardisation: Sections made of brain homogenate with spectrophotometrically measured enzyme activity were included in all incubation baths. In some cases the planned measurements were not attainable. This happened because of cutting-induced defects in brain slices or if the region was not obtainable due to the shift of the cut on the rostral-caudal axis.

## Statistical analysis
### Data imputation
(ref:LOT) An example of the imputed data in the nucleus of the lateral olfactory tract. We were missing 4 observations in total, 2 in each group. The figure shows the original data in the first column on the left and then 10 sets of data augmented with the imputed observations in magenta.

```{r lot-midrf, fig.cap="(ref:LOT)", fig.height=3.6, fig.width=5}
g_lot <- stripplot(mids_rf, LOT...1.3. ~ .imp, pch = 20, cex = 2, 
          xlab = "", ylab = "",
          scales = list(x = list(labels = c(0, paste0("i", 1:10)), cex = 1.2), 
                        y = list(cex = 1.2)), 
          par.settings = list(axis.line = list(col = 0)),
          panel=function(...){
            lims <- current.panel.limits()
            panel.stripplot(...)
            panel.abline(h=lims$ylim[1],v=lims$xlim[1], col = "black", lwd = 2)
          })
g_lot
```

Raw data contained COX measurements for 247 brain regions in 14 rats. Seven animals were from the control and 7 from the treatment group, respectively. For 69 brain regions, the data were complete. The rest of the brain regions were missing measurements for 1 or more animals. We decided to use multiple imputation of the missing data to make the best unbiased use of the available data. Because the original data had a significant number of highly correlated variables, we used random forest approach to deal with the multicollinearity. This model was applied to brain regions with 2 or fewer missing cases in each experimental condition. There were 190 such brain regions. The rest of the data were discarded. 10 multiply imputed datasets were generated by R-package *mice* [@R-mice]. In particular, the implementation of Breiman's [-@breimanClassificationRegressionTrees1984] random forest algorithm for R language was used as described by Doove and colleagues [-@dooveRecursivePartitioningMissing2014]. The algorithm was run for 20 iterations and showed good convergence. An example of the imputed data for one of the brain regions is shown on Figure \@ref(fig:lot-midrf).

### Testing for difference in mean COX levels
We applied Welch's two sample t-test to the 10 imputed datasets and used Barnard-Rubin's method to adjust the degrees of freedom for small samples in pooled data [@barnardMiscellaneaSmallsampleDegrees1999]. 

### Differential correlation analysis
Instead of comparing the levels of the enzyme in various brain regions between conditions, we can also study the inter-regional correlation patterns. In this study we employed the Differential Correlation Analysis (DCA) to identify the brain regions with different correlation profiles in control and ketamine group rats. Spearman rank-order correlation coefficients were calculated between all possible pairs of brain regions in each of the 10 multiply imputed data sets. The correlation coefficients were then converted to Fisher z-scores that follow normal distribution. Finally, the difference in z-scores between two conditions (`r knitr::asis_output("\u0394 z-score)")` was calculated and the significance of the difference score evaluated separately by 1000 permutation tests in each imputed data set. The empirical p-values from the permutation tests were then averaged across the 10 multiply imputed data sets. The medians of z-scores were calculated in a similar manner. Besides pairwise correlations, the median change in correlation for each brain region with all others (gives 1 summary score for each brain region) and the median change in correlation between two conditions across all brain regions (provides 1 summary score) were also calculated. The DCA procedures were carried with DGCA R software package [@mckenzieDGCAComprehensivePackage2016].  

### Cluster analysis
We performed cluster analysis on brain regions separately in control and ketamine groups. The data for cluster analysis was averaged over 10 multiply imputed datasets. Both Euclidian metric for raw COX measurements and Spearman metric for inter-regional correlation matrices were used to calculate the dissimilarity scores between brain regions. We used hierarchical clustering with Ward's D2 agglomeration to identify the clusters. k-medoids algorithm with Euclidian metric was also tried. The goodness of clustering was measured by gap statistic [@tibshiraniEstimatingNumberClusters2001] computed from 1000 Monte Carlo simulations. 

