
## Testing for difference in mean COX levels

```{r table-t-tests, cache=FALSE, results="asis"}
fp_table <- here::here("analysis", "data", "derived_data", "t_tests.rds")
# remove Cpu_background...1.3.
t_tests <- readr::read_rds(fp_table) %>% slice(-9) %>% dplyr::mutate(diff = paste(papaja::printnum(`mean difference`), '\u00B1', papaja::printnum(se))) %>% dplyr::select(-(5:8)) %>% dplyr::select(1, starts_with("diff"), everything()) 
t_tests$p <- papaja::printp(t_tests$p)
t_tests %>% 
kable(escape = F, col.names = linebreak(c("\nbrain region", "difference of the\nmeans \u00B1 SE", "\nt", "\ndf", "\np"), align = "c"), booktabs = T, format = "latex", caption = "Significant t-test results") %>% 
  kable_styling(bootstrap_options = "striped", font_size = 12)
```


After we have performed the Welch's two sample t-test on 10 multiply imputed datasets and pooled the estimates, the mean comparisons in `r nrow(t_tests)` brain regions emerged as statistically significant. The results of t-test comparisons are provided in table \@ref(tab:table-t-tests). In all significant comparisons ketamine group rats had higher levels of cytochrome c oxidase activity.

## Differential correlation analysis
```{r dca-overall-p, cache=FALSE, include=FALSE}
fp_overall_pvalue <- here::here("analysis", "data", "derived_data", "dgca_imputed_overall_pvalue.rds")
p_value <- readr::read_rds(fp_overall_pvalue) 
```

```{r dca-regional-average, cache=FALSE, results="asis"}
fp_avg_zscore <- here::here("analysis", "data", "derived_data", "dgca_imputed_ave_zscore.rds")
sign_cor_changes <- readr::read_rds(fp_avg_zscore)
sign_cor_changes$p <- papaja::printp(sign_cor_changes$p)
sign_cor_changes %>% dplyr::rename("brain region" = regions, "\u0394 z-score" = z_score_diff) %>% papaja::apa_table(escape = FALSE, caption = "Regional median changes in correlation")
```


The median correlation between the two conditions did not differ significantly p = `r papaja::printnum(p_value)`. The regional median changes in correlations are provided in table \@ref(tab:dca-regional-average). The five regions have significantly more negative median correlations with the rest of brain regions in ketamine-treated rats. The locus coeruleus instead showed a borderline significant increase in the median correlation.

```{r dca-pairwise, cache=FALSE, include=FALSE}
fp_pos_df <- here("analysis", "data", "derived_data", "control_pos_df.rds")
fp_neg_df <- here("analysis", "data", "derived_data", "control_neg_df.rds")
control_pos_df <- readr::read_rds(fp_pos_df)
control_neg_df <- readr::read_rds(fp_neg_df)
```

```{r table-positive, cache=FALSE, results="asis"}
control_pos_df$p <- papaja::printp(control_pos_df$p)
control_pos_df %>% dplyr::rename("region 1" = region1,
  "region 2" = region2, "\u0394 z-score" = z_score_diff, "r control" = control_cor, 
  "r ketamine" = treatment_cor) %>% 
  kable(booktabs = T, format = "latex", longtable = TRUE,
      caption = "Decreased pairwise correlations in ketamine rats") %>%
kable_styling(bootstrap_options = "striped", font_size = 12, latex_options = "repeat_header")
```

```{r table-negative, cache=FALSE, results="asis"}
control_neg_df$p <- papaja::printp(control_neg_df$p) 
control_neg_df %>% dplyr::rename("region 1" = region1,
  "region 2" = region2, "\u0394 z-score" = z_score_diff, "r control" = control_cor, 
  "r ketamine" = treatment_cor) %>% 
  kable(booktabs = T, format = "latex", longtable = TRUE,
      caption = "Increased pairwise correlations in ketamine rats") %>%
kable_styling(bootstrap_options = "striped", font_size = 12)
```

```{r pairwise-names, include=FALSE, cache=FALSE}
fp_negative_pairs <- here("analysis", "data", "derived_data", "negative_pairs.rds")
fp_positive_pairs <- here("analysis", "data", "derived_data", "positive_pairs.rds")
names_pos <- readr::read_rds(fp_positive_pairs)
names_neg <- readr::read_rds(fp_negative_pairs)
```

(ref:corplot) Graphical representation of changes in correlation coefficients after subchronic
ketamine treatment. Line segments connect pairs of brain regions with significant difference in correlation scores between the two conditions. The widths of line segments correspond to the magnitude of the difference between two correlation scores.

```{r cor-plot, fig.cap="(ref:corplot)", fig.height=8, fig.width=8}
g_graph_fp <- here("analysis", "data", "derived_data", "g_graph.rds")
g <- readr::read_rds(g_graph_fp)
cor_plot <- g + geom_edge_link(aes(color = direction, width = abs(z_score_diff)),
                 check_overlap = TRUE) +
  geom_node_point(aes(color = as.factor(nomenclature)), size = 3) +
  geom_node_text(aes(label = name), vjust = -0.5, hjust = 0.5, family = "serif") +
  scale_edge_color_brewer(palette = "Pastel1") +
  scale_color_brewer(palette = "Set1") + labs(
    title = "Changes in correlation scores after subchronic ketamine treatment",
    color = "Brain structures:",
    edge_color = "Change in pairwise correlations vs. control group:",
    edge_width = "\u0394 z-score:") +
  theme_void() + theme(legend.position = "bottom",
                       legend.box = "vertical",
                       plot.margin=unit(c(0.5,1,1,1.1),"cm"),
                       plot.title = element_text(hjust = 0.5, size = 12, face = "bold",
                                                 margin=margin(0,0,20,0)),
                       legend.text = element_text(size=rel(0.9)),
                       legend.key.width = unit(0.4, "cm"),
                       legend.key.height = unit(0.5, "cm")) +
  guides(edge_color = guide_legend(override.aes = list(edge_width = 3, linetype = 0)))
cor_plot
```

For the presentation of significantly changed pairwise correlation coefficients we chose brain regions that exhibited significant change in at least 2 pairs. In majority of the cases (`r nrow(control_pos_df)`), the positive association was abolished (high positive correlation either became non-significant or negative) in ketamine rats at p level below 0.01 (table \@ref(tab:table-positive)). Op(-5.8) was the brain region with the biggest number of significant pairwise changes in correlation (8 pairs). The other brain regions in table \@ref(tab:table-positive) had between 2 and 4 pairwise reductions in the correlation coefficients and were as follows `r sort(names_pos[-10]) %>%  stringi::stri_paste(., collapse = ", ")`. Conversely, the opposite pattern was observed in `r nrow(control_neg_df)` pairs, where the correlation increased after ketamine adminsitration (table \@ref(tab:table-negative)). Here, the brain regions with 2 pairwise reversals in correlation direction were as follows: `r sort(names_neg) %>%  stringi::stri_paste(., collapse = ", ")`.
The pairwise changes in correlation coefficients after subchronic administration of ketamine are summarised on figure \@ref(fig:cor-plot).

## Cluster analysis
We tried several approaches to clustering the data. Neither approach revealed a clear macrostructure in the data. The best fitting number of clusters for both control and ketamine rats was estimated to be 1.  
