## Testing for difference in mean COX levels

```{r table-t-tests, cache=FALSE, results="asis"}
fp_table <- here::here("analysis", "data", "derived_data", "t_tests.rds")
t_tests <- readr::read_rds(fp_table) 
t_tests$p <- papaja::printp(t_tests$p)
papaja::apa_table(t_tests, escape = TRUE, caption = "Significant t-test results")
```


After we have performed the Welch's two sample t-test on 10 multiply imputed datasets and pooled the estimates, the mean comparisons in 10 brain regions emerged as statistically significant. The 11th result was barely significant (GP background) and we decided to exclude it from the further analysis, as results in GP was already significantly different. The results of t-test are provided in Table \@ref(tab:table-t-tests). In all significant cases ketamine group rats had higher levels of cytochrome c oxidase activity.

## Differential correlation analysis
```{r dca-overall-p, cache=FALSE, include=FALSE}
fp_overall_pvalue <- here::here("analysis", "data", "derived_data", "dgca_imputed_overall_pvalue.rds")
p_value <- readr::read_rds(fp_overall_pvalue) 
```

```{r dca-regional-average, cache=FALSE, results="asis"}
fp_avg_zscore <- here::here("analysis", "data", "derived_data", "dgca_imputed_ave_zscore.rds")
sign_cor_changes <- readr::read_rds(fp_avg_zscore)
sign_cor_changes$p <- papaja::printp(sign_cor_changes$p)
sign_cor_changes %>% dplyr::rename("brain region" = regions, "delta z-score" = z_score_diff) %>% papaja::apa_table(escape = TRUE, caption = "Regional median changes in correlation")
```


The average correlation between the two conditions did not differ significantly p = `r papaja::printnum(p_value)`. The regional median changes in correlations are provided in table \@ref(tab:dca-regional-average). The five regions have significantly more negative median correlations with the rest of brain regions in ketamine-treated rats, Locus coeruleus instead showed a borderline significant increase in the median correlation.

```{r dca-pairwise, cache=FALSE, include=FALSE}
fp_pos_df <- here("analysis", "data", "derived_data", "control_pos_df.rds")
fp_neg_df <- here("analysis", "data", "derived_data", "control_neg_df.rds")
control_pos_df <- readr::read_rds(fp_pos_df)
control_neg_df <- readr::read_rds(fp_neg_df)
```

```{r table-positive, cache=FALSE, results="asis"}
control_pos_df$p <- papaja::printp(control_pos_df$p)
control_pos_df %>% dplyr::rename("region 1" = region1,
  "region 2" = region2, "delta z-score" = z_score_diff, "r control" = control_cor, 
  "r ketamine" = treatment_cor) %>% 
  kable(booktabs = T, format = "latex", longtable = TRUE,
      caption = "Decreased pairwise correlations in ketamine rats") %>%
kable_styling(bootstrap_options = "striped", font_size = 12, latex_options = "repeat_header")
```

```{r table-negative, cache=FALSE, results="asis"}
control_neg_df$p <- papaja::printp(control_neg_df$p) 
control_neg_df %>% dplyr::rename("region 1" = region1,
  "region 2" = region2, "delta z-score" = z_score_diff, "r control" = control_cor, 
  "r ketamine" = treatment_cor) %>% 
  kable(booktabs = T, format = "latex", longtable = TRUE,
      caption = "Increased pairwise correlations in ketamine rats") %>%
kable_styling(bootstrap_options = "striped", font_size = 12)
```
For the presentation of significantly changed pairwise correlation coefficients we chose brain regions that exhibited significant change in at least 2 pairs. 
As for pairwise comparisons between brain regions in `r nrow(control_pos_df)` cases the correlation has become more negative in ketamine rats at p level below 0.01 (table \@ref(tab:table-positive)). Op....5.8. was the brain region with the biggest number of  significant pairwise changes in correlation (8 pairs). The other brain regions in table \@ref(tab:table-positive) had between 2 and 4 pairwise reductions in the correlation coefficients. Conversely, in `r nrow(control_neg_df)` pairs the correlation was increased  after ketamine treatment (table \@ref(tab:table-negative)). 
