---
title             : "Cytochrome c oxidase metabolic mapping in a chronic ketamine administration model"
shorttitle        : "cytochrome oxidase after chronic ketamine"
author:
  - name          : "Denis Matrov"
    affiliation   : "1,3"
  - name          : "Sophie Imbeault"
    affiliation   : "2"
  - name          : "Margus Kanarik"
    affiliation   : "3"
  - name          : "Marianna Shkolnaya"
    affiliation   : "2"
  - name          : "Patricia Schikorra"
    affiliation   : "2"
  - name          : "Ergo Miljan"
    affiliation   : "2"
  - name          : "Ruth Shimmo"
    affiliation   : "2"
  - name          : "Jaanus Harro"
    affiliation   : "3"  
    corresponding : yes
    address       : "Postal address"
    email         : "jaanus.harro@ut.ee"

affiliation:
  - id            : "1"
    institution   : "Department of Neuroscience, Graduate School of Medicine, Kyoto University, Kyoto, Japan"
  - id            : "2" 
    institution   : "Tallinn University Centre of Excellence in Neural and Behavioural Sciences, School of Natural Sciences and Health, Tallinn University, Tallinn, Estonia" 
  - id            : "3"
    institution   :  "Division of Neuropsychopharmacology, Department of Psychology, University of Tartu, Tartu, Estonia"
    
keywords          : ["rats", "cytochrome oxidase", "ketamine"]
wordcount         : "X"

bibliography      : ["R-packages.bib", "references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
csl               : "../templates/behavioural-brain-research.csl"
output            : 
  papaja::apa6_pdf:
    latex_engine: xelatex

header-includes:
- \usepackage{booktabs}
- \usepackage{makecell}
---

```{r setup, echo = FALSE, include = FALSE, cache=FALSE}
library("papaja")
library("ggplot2")
library("tidyverse")
library("scales")
library("gridExtra")
library("mice")
library("here")
library("kableExtra")
library("ggraph")
library("grid")
library("igraph")

# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(
  cache = TRUE,
  cache.comments = FALSE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  echo = FALSE,
  strip.white = TRUE,
  comment = "#>",
  fig.path = "../figures/",
  results = "asis",
  tidy = "styler",
  dev = c('pdf', 'tiff'),
  dev.args = list(pdf = list(colormodel = 'cmyk'), 
                  tiff = list(compression = 'lzw')),
  dpi = 600,
  fig.width = 7,
  fig.height = 7,
  concordance = TRUE,
  global.par = TRUE
)

par(font.main = 1, cex.main = 1.05)
theme_set(theme_apa(base_size = 15) + theme(legend.position = "bottom"))
```



<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

# Introduction
The term "schizophrenia" was first used by Bleuler in reference to the separation of function between personality, thinking, memory, and perception of those afflicted with this disorder [@stotz-ingenlathEpistemologicalAspectsEugen2000]. Although prevalence is comparatively low (~0.7%), the direct and indirect costs are high [@chongGlobalEconomicBurden2016], making finding an effective treatment important to both patients and society. Schizophrenia symptoms are divided into three categories: positive (hallucinations, delusions), negative (anhedonia, blunted affect, asociality), and cognitive (impaired executive function and memory). The degree of heritability is high, up to 80% [@harrisonGenesSchizophreniaRecent2003], however the underlying genetics are complex [@harrisonSchizophreniaGenesGene2005]. Many of the genes identified in recent genome-wide association studies are implicated either directly or indirectly in glutamate signalling [@schizophreniaworkinggroupofthepsychiatricgenomicsconsortiumBiologicalInsights1082014; @stotz-ingenlathEpistemologicalAspectsEugen2000]. 

Glutamate is the most ubiquitous neurotransmitter in mammalian brain: about 60% of neurons contain glutamate, and virtually all neurons have some type of glutamate receptor. Early pharmacological observations of dissociative anaesthetics ketamine and phencyclidine (PCP) have lead to the hypothesis that glutamate plays a causal role in schizophrenia. These drugs affect the brain by blocking a major type of ionotropic glutamate receptor called N-methyl-D-aspartate receptor (NMDAR). Modulation of the postsynaptic NMDARs is crucial for activity-dependent synaptic plasticity and memory. By analogy, a hypofunction of NMDAR-mediated glutamatergic signalling was implicated in the pathophysiology of schizophrenia[@coyleGlutamateSchizophreniaDopamine2006]. Evidence for this hypothesis partly relies on the observation that when normal adults are administered an NMDAR antagonist, such as ketamine, they develop negative, cognitive and positive symptoms like those seen in schizophrenia [@adlerComparisonKetamineinducedThought1999; @krystalSubanestheticEffectsNoncompetitive1994; @newcomerKetamineinducedNMDAReceptor1999]. Administration of these compounds to someone with schizophrenia exacerbates their symptoms [@lahtiEffectsKetamineNormal2001]. Furthermore, stimulating NMDAR-mediated signalling using agonists of the glycine modulatory site has been effective in alleviating some of the symptoms of schizophrenia in clinical trials [@heresco-levyPlacebocontrolledTrialDcycloserine2002; @tsaiDserineAddedAntipsychotics1998]. Despite the early enthusiasm, the therapeutic efficacy of the first generation of antipsychotic drugs developed to augment glutamatergic neurotransmission via allosteric modulation of NMDAR was not great [@buchananCognitiveNegativeSymptoms2007; @iwataEffectsGlutamatePositive2015]. Active research on the role of glutamate in schizophrenia continues, some of recent strategies include targeting the metabotropic glutamate receptors 2, 3, and 5, as well as cholinergic receptors [@baluNMDAReceptorSchizophrenia2016].

Subchronic administration of the NMDAR antagonists such as ketamine, PCP, or dizocilpine (MK-801) readily replicates cognitive and negative aspects of schizophrenia in animal models. A daily treatment of 7 to 10 days is sufficient to impair declarative and working memory, as well as executive functions in rodents [@badoEffectsLowdoseDserine2011; @liPersistingCognitiveDeficits2011; @neillAnimalModelsCognitive2010]. Negative symptoms are also replicated by a reduction of non-aggressive behaviour in the social interaction test [@beckerKetamineinducedChangesRat2004].

Cytochrome c oxidase (COX; EC 1.9.3.1) is a rate-limiting enzyme in complex IV of the mitochondrial electron transport chain, where its activity mirrors the generation of ATP. COX activity is tightly coupled to neuronal activity and primarily reflects long-term post-synaptic energy expenditure [@wong-rileyBigenomicRegulationCytochrome2012]. Histological processing of brain tissue allows to measure COX levels at high resolution and in large number of brain regions. It provides a snapshot of mainly excitatory neuronal activity of the entire brain and is an excellent method to look for novel therapeutic targets. Histology remains the best way to measure neuronal activity in smaller subcortical brain regions. Our study is the first to evaluate the effect of ketamine administration on regional brain metabolic activity by means of histochemistry. Previously, another NMDAR antagonist PCP was studied after chronic administration of 28 days in rats. That study showed reduced regional COX activity, especially in basal ganglia and septum [@princeNormalizationCytochromecOxidase1997]. Experimental design similar to the current study was employed for enzymatic assay of COX activity in brain tissue homogenates. Ketamine was administered at slightly lower dose (25mg/kg for 7 days). Increased enzyme activity in complex IV was observed in striatum and hippocampus 1 to 6 hours after the last dose of ketamine [@deoliveiraBehavioralChangesMitochondrial2011]. 

The aim of this study was to identify brain regions with persistently different oxidative energy metabolism levels after subchronic ketamine administration in the established animal model of human schizophrenia.

# Materials and Methods

```{r include-methods, child = "02-methods.Rmd"}
```


# Results

```{r include-results, child = "03-results.Rmd"}
```


# Discussion
Median inter-regional correlation coefficient was similar between the control and treatment conditions. In agreement, cluster analysis suggested that in both conditions brain activity is similar on macro level. Therefore, our first conclusion is that ketamine administration evoked a limited number of local metabolic changes in brain regions and their pathways. 

The comparison of mean enzyme activity levels between control and treatment conditions identified 9 brain regions with elevated COX activity after subchronic ketamine treatment. Of them, 2 regions represented sensory thalamus, 2 basal ganglia, 3 cortical areas including 2 areas in the cingulate cortex, 1 hippocampal dentate gyrus, and 1 deep layer of the superior colliculus.

Differential correlation analysis is complimentary to the comparison of the mean COX levels. It accounts for the shape of the distribution of the results and allows to compare pairwise regional interactions between conditions. Four layers of the superior colliculus and one subdivision of the periaqueductal gray matter showed reduced positive correlations over all brain regions after subchronic ketamine treatment. The result for locus coeruleus was close to significance in the opposite direction. Layers of the superior colliculus were also prominently different between two rat groups in regional pairwise correlations of their COX levels. Their correlations with cortical and hippocampal brain regions were strongly positive in the control condition and tended to become negative after subchronic ketamine administration. DMPAG(-5.8), CA3(−4.8), and PR(-4.8) were three other brain regions with 3-4 pairwise significant reductions in positive correlations in the latter condition. 
The opposite pattern of changes in the pairwise correlation coefficients between two conditions was also observed, but in the fewer brain regions. None of these brain regions showed increased positive pairwise correlations after ketamine administration with more than 2 other brain regions. There was also little overlap between brain regions that exhibited more positive and more negative pairwise correlations in the ketamine group rats. Therefore, we can conclude that the reduction of the positive pairwise correlations after subchronic ketamine administration is a more common pattern and various layers of the superior colliculus are important nodes in this process.  

DMPAG(-5.8) is identified with avoidance behaviour in rodents [@borelliEffectsOvineCRF2008]. 
LPMR(-3.8) and superficial layers of superior colliculus (SuG(-5.8), Op(-5.8)) receive direct retinal projections and are important for visually guided behaviours and multisensory integration [@allenVisualInputMouse2016; @dragerTopographyVisualSomatosensory1976]. MGV(−5.8) is the primary auditory relay nucleus of the thalamus that projects to the primary auditory cortex [@ryugoDifferentialTelencephalicProjections1974]. Acute doses of ketamine produce visual and auditory deficits in healthy humans and rodents [eg @hillhouseEffectsNoncompetitiveNmethylDaspartate2015; @umbrichtKetamineinducedDeficitsAuditory2000]. Our experimental design had a 24 h washout period between the last ketamine injection and the animal sacrifice, hence the observed changes in COX activity probably have a longer-term effect on behaviour. 

While subchronic ketamine administration in rodents is used as an animal model of human schizophrenia, the metabolic processes in the brain are likely to be quite different. Schizophrenia is a chronic disease characterised by progressive facets of neurodegeneration, such as reductions in the cortical, hippocampal gray matter [@fusar-poliProgressiveBrainChanges2013; @glausierDendriticSpinePathology2013] and increased inflammation [@andersonImmunoinflammatoryOxidativeNitrosative2013]. 

Besides NMDAR, ketamine binds to dopamine D~2~, serotonin 5-HT~2~, and opioid receptors [@kapurNMDAReceptorAntagonists2002].

# Conclusion

# Abbreviations
## Brain regions
Brain regions are abbreviated according to the reference brain atlas [@paxinosRatBrainStereotaxic2007]. For each brain region the distance to bregma in millimetres according to the reference brain secion is provided in parentheses.
APT(-4.8), Anterior pretectal nucleus; AuD(-4.8), Auditory cortex, secondary, dorsal;
AuV(-4.8), Auditory cortex, secondary, ventral; B(-1.3), Nucleus basalis; CA1(-4.8), Hippocampal CA1(Cornu Ammonis area 1); CA2(-4.8), Hippocampal CA2; CA3(-4.8), Hippocampal CA3; Cg1,2(-0.3), Cingulate cortex, areas 1 & 2; CPu(-1.3), Caudate putamen; CPuDM(+0.2), Caudate putamen, dorsomedial; CPuV(+0.2), Caudate putamen, ventral; DG(-3.8), Hippocampal dentate gyrus; DG(-4.8), Hippocampal dentate gyrus; DLG(-4.8), Dorsal lateral geniculate; DMPAG(-5.8), Dorsomedial periaqueductal gray; DP(+2.2), Dorsal peduncular cortex; Dp(-5.8), Superior colliculi, deep gray/white layer; DRD(-8.0), Dorsal raphe, dorsal; Ect(-4.8), Ectorhinal cortex; GP(-1.3), Globus pallidus; InG(-5.8), Superior colliculi, intermediate gray layer; InWh(-5.8), Superior colliculi, intermediate white layer; LC(-10.04), Locus coeruleus; LP(-4.8), Lateral posterior nucleus; LPMR(-3.8), Lateral posterior thalamus, mediorostral; LSD(-0.3), Lateral septal nucleus, dorsal; M2(-0.3), Motor cortex, secondary; MCPO(-0.3), Magnocellular preoptic nucleus; MGV(-5.8), Medial geniculate, ventral; MM(-4.8), Medial mamillary, medial; Op(-5.8), Superior colliculi, optic nerve layer; Pir(-1.3), Piriform cortex; PR(-4.8), Prerubral field; PtA(-4.8), Parietal association cortex; RSA(-5.8), Retrosplenial cortex, agranular; RSG(-4.8), Retrosplenial cortex, granular; RSGa(-5.8), Retrosplenial cortex, granular; RSGb(-5.8), Retrosplenial cortex, granular; RTg(-8.0), Reticulotegmental nucleus; Shi(+0.2), Septohippocampal nucleus; SPTg(-8.0), Subpeduncular tegmental nucleus; SubD(-5.8), Subiculum, dorsal; SuG(-5.8), Superior colliculi, superficial gray layer; V1M(-8.0), Primary visual cortex, monocular; V2(-4.8), Visual cortex, secondary. 


## Other
COC, cytochrome c oxidase; DCA, differential correlation analysis; NMDAR, N-methyl-D-aspartate receptor; PCP, phencyclidine; SE, standard error; 

# Acknowledgements

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->

\pagebreak

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

\pagebreak

