---
title             : "Cytochrome c oxidase metabolic mapping in a chronic ketamine administration model"
shorttitle        : "cytochrome oxidase after chronic ketamine"
author:
  - name          : "Denis Matrov"
    affiliation   : "1,3"
  - name          : "Sophie Imbeault"
    affiliation   : "2"
  - name          : "Margus Kanarik"
    affiliation   : "3"
  - name          : "Marianna Shkolnaya"
    affiliation   : "2"
  - name          : "Patricia Schikorra"
    affiliation   : "2"
  - name          : "Ergo Miljan"
    affiliation   : "2"
  - name          : "Ruth Shimmo"
    affiliation   : "2"
  - name          : "Jaanus Harro"
    affiliation   : "3"  
    corresponding : yes
    address       : "Ravila 14A, Chemicum, 50411, Tartu, Estonia"
    email         : "jaanus.harro@ut.ee"

affiliation:
  - id            : "1"
    institution   : "Department of Neuroscience, Graduate School of Medicine, Kyoto University, Kyoto, Japan"
  - id            : "2" 
    institution   : "Tallinn University Centre of Excellence in Neural and Behavioural Sciences, School of Natural Sciences and Health, Tallinn University, Tallinn, Estonia" 
  - id            : "3"
    institution   :  "Division of Neuropsychopharmacology, Department of Psychology, University of Tartu, Tartu, Estonia"
    
keywords          : ["rats", "cytochrome oxidase", "ketamine"]
wordcount         : "X"

bibliography      : ["R-packages.bib", "references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
csl               : "../templates/behavioural-brain-research.csl"
output            : 
  papaja::apa6_pdf:
    latex_engine: xelatex
  # papaja::apa6_word
header-includes:
- \usepackage{booktabs}
- \usepackage{makecell}
---

```{r setup, echo = FALSE, include = FALSE, cache=FALSE}
library("papaja")
library("ggplot2")
library("tidyverse")
library("scales")
library("gridExtra")
library("mice")
library("here")
library("kableExtra")
library("ggraph")
library("grid")
library("igraph")

# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(
  cache = TRUE,
  cache.comments = FALSE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  echo = FALSE,
  strip.white = TRUE,
  comment = "#>",
  fig.path = "../figures/",
  results = "asis",
  tidy = "styler",
  dev = c('pdf', 'tiff'),
  dev.args = list(pdf = list(colormodel = 'cmyk'), 
                  tiff = list(compression = 'lzw')),
  dpi = 600,
  fig.width = 7,
  fig.height = 7,
  concordance = TRUE,
  global.par = TRUE
)

par(font.main = 1, cex.main = 1.05)
theme_set(theme_apa(base_size = 15) + theme(legend.position = "bottom"))
```



<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

# Introduction
The term "schizophrenia" was first used by Bleuler in reference to the separation of function between personality, thinking, memory, and perception of those afflicted with this disorder [@stotz-ingenlathEpistemologicalAspectsEugen2000]. Although prevalence is comparatively low (~0.7%), the direct and indirect costs are high [@chongGlobalEconomicBurden2016], making finding an effective treatment important to both patients and society. Schizophrenia symptoms are divided into three categories: positive (hallucinations, delusions), negative (anhedonia, blunted affect, asociality), and cognitive (impaired executive function and memory). The degree of heritability is high, up to 80% [@harrisonGenesSchizophreniaRecent2003], however the underlying genetics are complex [@harrisonSchizophreniaGenesGene2005]. Tradionally, aetiology of schizophrenia was linked to the hyperactivity in the dopaminergic signal transduction. More recent advances in pharmacology and genetics indicate, however, that glutamatergic neurotransmission may play an important role. For example, many of the genes identified in recent genome-wide association studies are implicated either directly or indirectly in the glutamate signalling [@schizophreniaworkinggroupofthepsychiatricgenomicsconsortiumBiologicalInsights1082014; @stotz-ingenlathEpistemologicalAspectsEugen2000]. 

Glutamate is the most ubiquitous neurotransmitter in mammalian brain: about 60% of neurons contain glutamate, and virtually all neurons have some type of glutamate receptor. Early pharmacological observations of dissociative anaesthetics ketamine and phencyclidine (PCP) have lead to the hypothesis that glutamate plays a causal role in schizophrenia. These drugs affect the brain by blocking a major type of ionotropic glutamate receptor called N-methyl-D-aspartate receptor (NMDAR). Modulation of the postsynaptic NMDARs is crucial for activity-dependent synaptic plasticity and memory. By analogy, a hypofunction of NMDAR-mediated glutamatergic signalling was implicated in the pathophysiology of schizophrenia [@coyleGlutamateSchizophreniaDopamine2006]. Evidence for this hypothesis partly relies on the observation that when normal adults are administered an NMDAR antagonist, such as ketamine, they develop negative, cognitive and positive symptoms like those seen in schizophrenia [@adlerComparisonKetamineinducedThought1999; @krystalSubanestheticEffectsNoncompetitive1994; @newcomerKetamineinducedNMDAReceptor1999]. Administration of these compounds to someone with schizophrenia exacerbates their symptoms [@lahtiEffectsKetamineNormal2001]. Furthermore, stimulating NMDAR-mediated signalling using agonists of the glycine modulatory site has been effective in alleviating some of the symptoms of schizophrenia in clinical trials [@heresco-levyPlacebocontrolledTrialDcycloserine2002; @tsaiDserineAddedAntipsychotics1998]. Despite the early enthusiasm, the therapeutic efficacy of the first generation of antipsychotic drugs developed to augment glutamatergic neurotransmission via allosteric modulation of NMDAR was found to be rather limited [@buchananCognitiveNegativeSymptoms2007; @iwataEffectsGlutamatePositive2015]. Glutamatergic neurotransmission is very complex, as it involves different receptor types, molecular adaptation mechanisms, and tight coupling between oxidative metabolism and glutamate-glutamine neurotransmitter cycle [@napolitanoNeurometabolicProfilingKetamine2016]. Therefore the nature of glutamatergic deficits and best ways to counter them remain very active areas of research [@baluNMDAReceptorSchizophrenia2016].

Subchronic administration of the NMDAR antagonists such as ketamine, PCP, or dizocilpine (MK-801) readily replicates cognitive and negative aspects of schizophrenia in animal models. A daily treatment of 7 to 10 days is sufficient to impair declarative and working memory, as well as executive functions in rodents [@badoEffectsLowdoseDserine2011; @liPersistingCognitiveDeficits2011; @neillAnimalModelsCognitive2010]. Negative symptoms are also replicated by a reduction of non-aggressive behaviour in the social interaction test [@beckerKetamineinducedChangesRat2004].

Cytochrome c oxidase (COX; EC 1.9.3.1) is a rate-limiting enzyme in complex IV of the mitochondrial electron transport chain, where its activity mirrors the generation of ATP. COX activity is tightly coupled to neuronal activity at molecular level and primarily reflects long-term post-synaptic energy expenditure [@wong-rileyBigenomicRegulationCytochrome2012]. 
Nuclear respiratory factor 1 (NRF-1) binds to COX subunit genes and functionally regulates neuronal metabolism. NRF-1 also co‐regulates AMPA glutamate receptor subunit 2 and NMDA receptor subunits 1 (NR1) and 2b (NR2b) genes, therefore mitochondrial energy generation and glutamate cell transduction are controlled by overlapping transcriptional mechanisms [@dharCouplingEnergyMetabolism2009; @dharNuclearRespiratoryFactor2009]. Histological processing of brain tissue allows to measure COX levels at high resolution and in large number of brain regions. It provides a snapshot of mainly excitatory neuronal activity of the entire brain and is an excellent method to look for novel therapeutic targets. Histology remains the best way to measure neuronal activity in smaller subcortical brain regions. Our study is the first to evaluate the effect of ketamine administration on regional brain metabolic activity by means of histochemistry. Previously, another NMDAR antagonist PCP was studied after chronic administration of 28 days in rats. That study showed reduced regional COX activity, especially in basal ganglia and septum [@princeNormalizationCytochromecOxidase1997]. Experimental design similar to the current study was employed for enzymatic assay of COX activity in brain tissue homogenates. Ketamine was administered at slightly lower dose (25 mg/kg for 7 days). Increased enzyme activity in complex IV was observed in striatum and hippocampus 1 to 6 h after the last dose of ketamine [@deoliveiraBehavioralChangesMitochondrial2011]. 

The aim of this study was to identify brain regions with persistently different oxidative energy metabolism levels after subchronic ketamine administration in the established animal model of human schizophrenia.

# Materials and Methods

```{r include-methods, child = "02-methods.Rmd"}
```


# Results

```{r include-results, child = "03-results.Rmd"}
```


# Discussion
Median inter-regional correlation coefficient was similar between the control and treatment conditions. In agreement, cluster analysis suggested that in both conditions brain activity is similar on macro level. Therefore, our first conclusion is that ketamine administration evoked a limited number of local metabolic changes in brain regions and their pathways. 

The comparison of mean enzyme activity levels between control and treatment conditions identified 9 brain regions with elevated COX activity after subchronic ketamine treatment. Of them, 2 regions represented sensory thalamus, 2 basal ganglia, 3 cortical areas (2 areas in the cingulate cortex and 1 in olfactory cortex), 1 hippocampal dentate gyrus, and 1 deep layer of the superior colliculus.

Differential correlation analysis is complimentary to the comparison of the mean COX levels. It accounts for the shape of the distribution of the results and allows to compare pairwise regional interactions between conditions. Four layers of the superior colliculus and one subdivision of the periaqueductal gray matter showed reduced positive correlations over all brain regions after subchronic ketamine treatment. The result for locus coeruleus was close to significance in the opposite direction. Layers of the superior colliculus were also prominently different between two rat groups in regional pairwise correlations of their COX levels. Their correlations with cortical and hippocampal brain regions were strongly positive in the control condition and tended to become negative after subchronic ketamine administration. DMPAG(-5.8), CA3(−4.8), and PR(-4.8) were three other brain regions with 3-4 pairwise significant reductions in positive correlations in the latter condition. 
The opposite pattern of changes in the pairwise correlation coefficients between two conditions was also observed, but in the fewer brain regions. None of these brain regions showed increased positive pairwise correlations after ketamine administration with more than 2 other brain regions. There was also little overlap between brain regions that exhibited more positive and more negative pairwise correlations in the ketamine group rats. Therefore, we can conclude that the reduction of the positive pairwise correlations after subchronic ketamine administration is a more common pattern and various layers of the superior colliculus are prominent nodes in this process.

Antagonosts of NMDA receptors, also called dissociative anaesthetics have a long history. 
PCP was developed first, but due to its toxicity is mainly famous as an illicit street drug. Ketamine is a safer PCP derivative, first synthesised in 1962 [@dominoPharmacologicEffectsCI5811965]. Dizocilpine was first described 20 years after ketamine [@clineschmidtAnticonvulsantActivity5methyl101982]. All three NMDA antagonists have been studied in animal models of schizophrenia, so while our main emphasis is on ketamine, we will briefly mention some of the studies with PCP and dizocilpine as well. Acute and subchronic administration of ketamine in subanaesthetic doses somewhat paradoxically elevates neuronal activity and neurotransmitter release in various brain regions. This effect has been recorded in both primates and rodents using several methods. For example, in humans, subanaesthetic infusion of ketamine during PET scan increased regional glucose metabolic rate in most brain regions, including the anterior and posterior cingulate cortex, caudate, and thalamus. No regions with decreased glucose metabolism were found [@langsjoEffectsSubanestheticKetamine2004]. Similarly, studies of brain haemodynamic activity in rats, monkeys, and humans all have shown quite consistent pattern of regional increases in the BOLD signal indicative of neuronal excitation [reviewed in @maltbieKetaminePharmacologicalImaging2017]. In all regions, peak BOLD response occurs 3–5 min after the start of ketamine infusion, which correlates well with the timecourse of ketamine peak blood levels [@deakinGlutamateNeuralBasis2008]. In monkeys, acute ketamine injection increases both spontaneous and task-related firing of layer V pyramidal cells in the prefrontal cortex [@wangNMDAReceptorsSubserve2013]. There is experimental evidence that while the activity of the pyramidal cells in increased after ketamine infusion, the inhibition of NMDA receptors mainly decreases the activity of putative GABA interneurons [@homayounNMDAReceptorHypofunction2007]. Therefore the at least some of the elevated cerebral excitation stems from disinhibition projection neurons that were released from the GABA-ergic inhibition. 

The effects of ketamine on neurotransmitter release are reviewed in [@napolitanoNeurometabolicProfilingKetamine2016]. Subchronic ketamine administration in rats releases glutamate in the frontal lobe [@chatterjeeNeurochemicalMolecularCharacterization2012; @kimVivoExVivo2011]. Acute subanaesthetic doses of ketamine (10-30 mg/kg) increased glutamate release in the prefrontal cortex of rats in a dose-dependent manner and required highest dose at 30 mg/kg to also elevate the outflow of dopamine and GABA there [@moghaddamActivationGlutamatergicNeurotransmission1997]. Similarly, acute and subchronic PCP administration increased glutamate release and reduced levels of GABA neuronal markers parvalbumin and GAD67 in the prefrontal cortex [@amitaiRepeatedPhencyclidineAdministration2012]. Acute dizocilpine administration increased dopamine and serotonin turnover in the frontal cortex, piriform cortex, hippocampus, and striatum [@hiramatsuComparisonBehavioralBiochemical1989; @loscherNmethylDaspartateReceptorAntagonist1991]. Only serotonin turnover was elevated in the thalamus after acute dose of dizocilpine [@loscherNmethylDaspartateReceptorAntagonist1991]. Similar results have been obtained with PCP [@hondoEffectPhencyclidineDopamine1994; @nabeshimaSerotonergicInvolvementPhencyclidineinduced1984]. Increased cerebral serotonin turnover after were also observed after acute and subchronic ketamine administration [@chatterjeeNeurochemicalMolecularCharacterization2012; @vargiuPossibleRoleBrain1978].
Chatterjee and colleagues [-@chatterjeeNeurochemicalMolecularCharacterization2012] have performed comprehensive research on acute and subchronic (10 days of treatment) effects of ketamine in mice. They found higher cortical, striatal, and hippocampal dopamine and acetylcholine turnover [@chatterjeeNeurochemicalMolecularCharacterization2012; see also @kikuchiEffectsKetaminePentobarbitone1997]. D~1~ and D~2~ dopamine, as well as 5-HT~1A~ receptor expressions in the cortex was significantly elevated. Glycine levels were reduced in cortex, striatum, and hippocampus after subchronic ketamine administration and only the cortical GABA levels were elevated. The gene-expression of two NMDAR subunits NR1 and NR2b was likewise higher in the cortex after subchronic ketamine administration [@chatterjeeNeurochemicalMolecularCharacterization2012]. Previously, the same group found that behavioural effects of subchronic ketamine regimen in mice persist for at least 10 days after the withdrawal of ketamine administration [@chatterjeeEffectChronicAcute2011]. Besides NMDAR, ketamine binds to dopamine D~2~ receptors in high affinity state, serotonin 5-HT~2A~, sigma 1, and opioid receptors [@frohlichReviewingKetamineModel2014; @kapurNMDAReceptorAntagonists2002]. In summary, ketamine binds to numerous types of receptors and its application rapidly reconfigures and elevates neuronal activity.
Finally, in some studies the metabolic activity was also measured from histological preparations similar to the current study. A quantitative autoradiographic ^14^C-2-deoxyglucose (2-DG) technique is used to measure the glucose oxidation just minutes after the acute dose of NMDAR antaganist. Pronounced increases in local cerebral
glucose utilisation were observed after acute dizocilpine dose in the olfactory areas and in a number of brain areas in the limbic system (including dentate gyrus, posterior cingulate cortex, and mammillary body) [@kurumajiEffectsMK801Local1989]. Similar results were found for ketamine (35 mg/kg). Increased 2-DG uptake was recorded in cingulate, retrosplenial, and piriform cortices, dentate gyrus, various thalamic nuclei, although there was no change in medial geniculate [@duncanMetabolicMappingRat1998]. While in pharmacological studies PCP was quite similar to ketamine, on histological sections the reported results are opposite.
Chronic and subchronic treatments of PCP were associated with reduced glucose metabolism in some prefrontal areas in the cortex and auditory centres of the brain [@cochranInductionMetabolicHypofunction2003]. Reduced brain metabolism after the chronic PCP regimen was also identified in COX activity, but mainly in basal ganglia and septum [@princeNormalizationCytochromecOxidase1997]. Our results show elevated metabolic activity in 9 brain regions in total while we did not find any brain region with reduced COX levels. Such results are well in agreement with previous studies and show that subchronic ketamine regimen potentiates neural activity. Our experimental design had a 24 h washout period between the last ketamine injection and the animal sacrifice, hence the observed changes in COX activity likely reflect persistent neurochemical and behavioural changes.

TO DO:

In the rest of the text the control group is used as reference. The differential results we discuss apply to ketamine group rats when not explicitly specified otherwise. 
DMPAG(-5.8) is identified with avoidance behaviour in rodents [@borelliEffectsOvineCRF2008]. 
LPMR(-3.8) and superficial layers of superior colliculus (SuG(-5.8), Op(-5.8)) receive direct retinal projections and are important for visually guided behaviours and multisensory integration [@allenVisualInputMouse2016; @dragerTopographyVisualSomatosensory1976]. COX activity in MGV(−5.8) was elevated in absolute levels. Its correlations with Dp(-5.8) and DRD(−8.0) has also become significanlty positive. MGV(−5.8) is the primary auditory relay nucleus of the thalamus that projects to the primary auditory cortex [@ryugoDifferentialTelencephalicProjections1974]. 

Acute doses of ketamine produce visual and auditory deficits in healthy humans and rodents [eg @hillhouseEffectsNoncompetitiveNmethylDaspartate2015; @umbrichtKetamineinducedDeficitsAuditory2000]. 

While subchronic ketamine administration in rodents is used as an animal model of human schizophrenia, the metabolic processes in the brain are likely to be quite different. Schizophrenia is a chronic disease characterised by progressive facets of neurodegeneration, such as reductions in the cortical, hippocampal gray matter [@fusar-poliProgressiveBrainChanges2013; @glausierDendriticSpinePathology2013] and increased inflammation [@andersonImmunoinflammatoryOxidativeNitrosative2013]. 

Pathological changes in the form of vacuolization of neuronal cytoplasm called Olney's lesions after their discoverer [@olneyPathologicalChangesInduced1989] appear rapidly already 2 h after acute dose of ketamine, PCP, and dizocilpine in posterior and retrosplenial cingulate cortex. Subchronic administration shows lack of the cumulative pathology and there is even dissipation of pahtological changes. 


> Schizophrenia is characterised by abnormal oscillatory activity, particularly in the gamma band (30–80 Hz), where both reduced and elevated gamma synchrony were found dependent on the task and brain region. Gamma rhythms are associated with primary sensory representation.
Visual-evoked gamma band oscillations have lower frequency in schizophrenia patients (Spencer et al., 2004), as well as reduced phase locking (Spencer et al., 2008), decreased interhemispheric phase coherence (in contradistinction to the increased anterior-posterior phase coherence seen in healthy controls), and increased latency to coherence changes (Spencer et al., 2003). 


> auditory steady-state electroencephalogram responses are impaired during the first episode of psychosis [@spencerGammabandAuditorySteadystate2008].  SZ exhibit reduced γ-band neural synchrony during the perception of visual Gestalt patterns [@spencerNeuralSynchronyIndexes2004].

The limitation of our study was relatively low number of the animals, therefore some significant effects of ketamine administartion are lilely to be masked by the presence of 1 or 2 outliers.

# Conclusion

# Abbreviations
## Brain regions
Brain regions are abbreviated according to the reference brain atlas [@paxinosRatBrainStereotaxic2007]. For each brain region the distance to bregma in millimetres according to the reference brain secion is provided in parentheses.
APT (-4.8), Anterior pretectal nucleus; AuD (-4.8), Auditory cortex, secondary, dorsal;
AuV (-4.8), Auditory cortex, secondary, ventral; B (-1.3), Nucleus basalis; CA1 (-4.8), Hippocampal CA1(Cornu Ammonis area 1); CA2 (-4.8), Hippocampal CA2; CA3 (-4.8), Hippocampal CA3; Cg1,2 (-0.3), Cingulate cortex, areas 1 & 2; CPu (-1.3), Caudate putamen; CPuDM (+0.2), Caudate putamen, dorsomedial; CPuV (+0.2), Caudate putamen, ventral; DG (-3.8), Hippocampal dentate gyrus; DG (-4.8), Hippocampal dentate gyrus; DLG (-4.8), Dorsal lateral geniculate; DMPAG (-5.8), Dorsomedial periaqueductal gray; DP (+2.2), Dorsal peduncular cortex; Dp (-5.8), Superior colliculi, deep gray/white layer; DRD (-8.0), Dorsal raphe, dorsal; Ect (-4.8), Ectorhinal cortex; GP (-1.3), Globus pallidus; InG (-5.8), Superior colliculi, intermediate gray layer; InWh (-5.8), Superior colliculi, intermediate white layer; LC (-10.04), Locus coeruleus; LP (-4.8), Lateral posterior nucleus; LPMR (-3.8), Lateral posterior thalamus, mediorostral; LSD (-0.3), Lateral septal nucleus, dorsal; M2 (-0.3), Motor cortex, secondary; MCPO (-0.3), Magnocellular preoptic nucleus; MGV (-5.8), Medial geniculate, ventral; MM (-4.8), Medial mamillary, medial; Op (-5.8), Superior colliculi, optic nerve layer; Pir (-1.3), Piriform cortex; PR (-4.8), Prerubral field; PtA (-4.8), Parietal association cortex; RSA (-5.8), Retrosplenial cortex, agranular; RSG (-4.8), Retrosplenial cortex, granular; RSGa (-5.8), Retrosplenial cortex, granular; RSGb (-5.8), Retrosplenial cortex, granular; RTg (-8.0), Reticulotegmental nucleus; Shi (+0.2), Septohippocampal nucleus; SPTg (-8.0), Subpeduncular tegmental nucleus; SubD (-5.8), Subiculum, dorsal; SuG (-5.8), Superior colliculi, superficial gray layer; V1M (-8.0), Primary visual cortex, monocular; V2 (-4.8), Visual cortex, secondary. 


## Other
2-DG, ^14^C-2-deoxyglucose; 5-HIAA, 5-hydroxyindoleacetic acid; AMPA, $\alpha$-amino-3-hydroxy-5-methylisoxazole-4-propionate; BOLD, blood oxygen level dependent;
COX, cytochrome c oxidase; DCA, differential correlation analysis; GABA, $\gamma$-aminobutyric acid; GAD67, glutamic acid decarboxylase-67; NMDAR, N-methyl-D-aspartate receptor; NRF-1, Nuclear respiratory factor 1; PCP, phencyclidine; PET, positron emission tomography; SE, standard error; 

# Acknowledgements

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->

\pagebreak

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

\pagebreak

