---
title             : "Cytochrome c oxidase metabolic mapping in a subchronic ketamine administration model"
shorttitle        : "Oxidative metabolism after subchronic ketamine"
author:
  - name          : "Denis Matrov"
    affiliation   : "1,3,*"
  - name          : "Sophie Imbeault"
    affiliation   : "2,*"
  - name          : "Margus Kanarik"
    affiliation   : "3"
  - name          : "Marianna Shkolnaya"
    affiliation   : "2"
  - name          : "Patricia Schikorra"
    affiliation   : "2"
  - name          : "Ergo Miljan"
    affiliation   : "2"
  - name          : "Ruth Shimmo"
    affiliation   : "2"
  - name          : "Jaanus Harro"
    affiliation   : "3"  
    corresponding : yes
    address       : "Ravila 14A, Chemicum, 50411, Tartu, Estonia"
    email         : "jaanus.harro@ut.ee"

affiliation:
  - id            : "1"
    institution   : "Department of Neuroscience, Graduate School of Medicine, Kyoto University, Kyoto, Japan"
  - id            : "2" 
    institution   : "Tallinn University Centre of Excellence in Neural and Behavioural Sciences, School of Natural Sciences and Health, Tallinn University, Tallinn, Estonia" 
  - id            : "3"
    institution   :  "Division of Neuropsychopharmacology, Department of Psychology, University of Tartu, Tartu, Estonia"
  - id            : "*"
    institution   :  "Both authors contributed equally to this manuscript"
    
keywords          : ["cytochrome oxidase", "histochemistry", "ketamine", "rats"]
wordcount         : "X"

bibliography      : ["R-packages.bib", "references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
csl               : "../templates/behavioural-brain-research.csl"
output            : 
  papaja::apa6_pdf:
    latex_engine: xelatex
  # papaja::apa6_word
header-includes:
- \usepackage{booktabs}
- \usepackage{makecell}
---

```{r setup, echo = FALSE, include = FALSE, cache=FALSE}
library("papaja")
library("ggplot2")
library("tidyverse")
library("scales")
library("gridExtra")
library("mice")
library("here")
library("kableExtra")
library("ggraph")
library("grid")
library("igraph")

# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(
  cache = TRUE,
  cache.comments = FALSE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  echo = FALSE,
  strip.white = TRUE,
  comment = "#>",
  fig.path = "../figures/",
  results = "asis",
  tidy = "styler",
  dev = c('pdf', 'tiff'),
  dev.args = list(pdf = list(colormodel = 'cmyk'), 
                  tiff = list(compression = 'lzw')),
  dpi = 600,
  fig.width = 7,
  fig.height = 7,
  concordance = TRUE,
  global.par = TRUE
)

par(font.main = 1, cex.main = 1.05)
theme_set(theme_apa(base_size = 15) + theme(legend.position = "bottom"))
```



<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

# Introduction
Although the lifetime prevalence of schizophrenia is comparatively low (0.3-0.7%) [@americanpsychiatricassociationDiagnosticStatisticalManual2013], the direct and indirect costs are high owing to early onset and disabling chronic course [@chongGlobalEconomicBurden2016], making finding an effective treatment important to both patients and society. Schizophrenia symptoms are divided into three categories: positive (hallucinations, delusions), negative (anhedonia, blunted affect, social withdrawal), and cognitive (impaired executive function and memory). Traditionally, aetiology of schizophrenia has been linked to the hyperactivity in the dopaminergic signal transduction. However, early research on the behavioural effects of glutamate NMDA receptor antagonists phencyclidine (PCP) and ketamine and more recent advances in neurobiology and genetics have indicated that glutamate signal transduction also plays an important role [@javittGlutamatergicTheoriesSchizophrenia2010, @schizophreniaworkinggroupofthepsychiatricgenomicsconsortiumBiologicalInsights1082014].

Glutamate is the most ubiquitous neurotransmitter in mammalian brain: about 60% of neurons contain glutamate, and virtually all neurons have some type of glutamate receptor. Early pharmacological observations of dissociative anaesthetics ketamine and PCP have lead to the hypothesis that glutamate plays a causal role in schizophrenia. These drugs inactivate signal transduction by blocking the ion channel within a major type of ionotropic glutamate receptor called N-methyl-D-aspartate receptor (NMDAR). Modulation of the postsynaptic NMDARs is crucial for activity-dependent synaptic plasticity and memory. By analogy, a hypofunction of NMDAR-mediated glutamatergic signalling was implicated in the pathophysiology of schizophrenia [@coyleGlutamateSchizophreniaDopamine2006]. Evidence for this hypothesis partly relies on the observation that when normal adults are administered an NMDAR antagonist, such as ketamine, they develop negative, cognitive and positive symptoms like those seen in schizophrenia [@adlerComparisonKetamineinducedThought1999; @krystalSubanestheticEffectsNoncompetitive1994; @newcomerKetamineinducedNMDAReceptor1999]. Administration of these compounds to someone with schizophrenia exacerbates their symptoms [@lahtiEffectsKetamineNormal2001]. Furthermore, stimulating NMDAR-mediated signalling using agonists of the glycine modulatory site has been effective in alleviating some of the symptoms of schizophrenia in clinical trials [@heresco-levyPlacebocontrolledTrialDcycloserine2002; @tsaiDserineAddedAntipsychotics1998]. Despite the early enthusiasm, the therapeutic efficacy of the first antipsychotic drugs developed to augment glutamatergic neurotransmission via allosteric modulation of NMDAR was found to be rather limited [@buchananCognitiveNegativeSymptoms2007; @iwataEffectsGlutamatePositive2015]. Glutamatergic neurotransmission is very complex, as it involves different receptor types, molecular adaptation mechanisms, and tight coupling between oxidative metabolism and glutamate-glutamine neurotransmitter cycle [@napolitanoNeurometabolicProfilingKetamine2016]. Therefore the nature of glutamatergic deficits in schizophrenia and best ways to counter them remain very active areas of research [@baluNMDAReceptorSchizophrenia2016].

Subchronic administration of the NMDAR antagonists such as ketamine, PCP, or dizocilpine (MK-801) readily replicates cognitive and negative aspects of schizophrenia in animal models. A daily treatment of 7 to 10 days is sufficient to impair declarative and working memory, as well as executive functions in rodents [@badoEffectsLowdoseDserine2011; @liPersistingCognitiveDeficits2011; @neillAnimalModelsCognitive2010]. Negative symptoms are also replicated by a reduction of non-aggressive behaviour in the social interaction test [@beckerKetamineinducedChangesRat2004].

Cytochrome c oxidase (COX; EC 1.9.3.1) is a key enzyme in complex IV of the mitochondrial electron transport chain, where its activity mirrors the generation of ATP. COX activity is tightly coupled to neuronal activity at molecular level and primarily reflects long-term post-synaptic energy expenditure [@wong-rileyBigenomicRegulationCytochrome2012]. 
Nuclear respiratory factor 1 (NRF-1) binds to COX subunit genes and functionally regulates neuronal metabolism. NRF-1 also co‐regulates AMPA glutamate receptor subunit 2 and NMDA receptor subunits 1 (NR1) and 2b (NR2b) genes, therefore mitochondrial energy generation and glutamate cell transduction are controlled by overlapping transcriptional mechanisms [@dharCouplingEnergyMetabolism2009; @dharNuclearRespiratoryFactor2009]. Histological processing of brain tissue allows to measure COX levels at high resolution and in large number of brain regions. It provides a snapshot of mainly excitatory neuronal activity of the entire brain and is an excellent method to look for novel biochemical targets. 
COX activity has been found to be increased in post-mortem brain samples of schizophrenia patients, and in correlation with their intellectual and emotional impairment [@princeMitochondrialFunctionDifferentially1999; @princePutamenMitochondrialEnergy2000]. Human patients usually have a history of chronic antipsychotic drug treatment and changes of COX activity also reflect the effects of medication. Accordingly, COX activity in a number of brain regions was enhanced by the administration of several antipsychotic drugs in rats [@princeHistochemicalDemonstrationAltered1998]. In this regard, animal models allow more precision in disentangling the brain localisations of dysfunctions and therapeutic effect of different drugs. The data on brain metabolic effects of subanaesthetic treatment with NMDAR antagonists is relatively sparse. One study assessed the effect of PCP after chronic administration of 28 days in rats and reported reduced COX activity in several brain regions, especially in basal ganglia and septum [@princeNormalizationCytochromecOxidase1997]. In another study employing the enzymatic assay of COX activity in brain tissue homogenates after ketamine had been administered at 25 mg/kg for 7 days, increased enzyme activity was observed in striatum and hippocampus 1 to 6 h after the last dose of ketamine  [@deoliveiraBehavioralChangesMitochondrial2011]. 

The aim of this study was to identify brain regions with a) persistently different oxidative energy metabolism levels by comprehensive mapping and b) to reveal potentially differential regional activity co-variance after subchronic ketamine administration.

# Materials and Methods

```{r include-methods, child = "02-methods.Rmd"}
```


# Results

```{r include-results, child = "03-results.Rmd"}
```


# Discussion
The comparison of mean enzyme activity levels between control and treatment conditions showed an increase in the oxidative metabolic activity in several brain regions. Of them, 2 regions represented sensory thalamus, 2 basal ganglia, 3 cortical areas (2 areas in the cingulate cortex and 1 in olfactory cortex), 1 hippocampal dentate gyrus, and 1 deep layer of the superior colliculi. Results from current study are in good agreement with previous studies and show that subchronic ketamine regimen potentiates neural activity. Our experimental design had a 24 h washout period between the last ketamine injection and the animal sacrifice, hence the observed changes in COX activity likely reflect persistent neurochemical and behavioural adaptations.

Antagonists of NMDA receptors, also called dissociative anaesthetics have a long history. 
PCP was developed first, but due to its toxicity is mainly famous as an illicit street drug. Ketamine is a safer PCP derivative, first synthesised in 1962 [@dominoPharmacologicEffectsCI5811965]. Dizocilpine was first described 20 years after ketamine [@clineschmidtAnticonvulsantActivity5methyl101982]. All three NMDA antagonists have been studied in animal models of schizophrenia, so while our main emphasis is on ketamine, we will briefly mention some of the studies with PCP and dizocilpine as well. Acute and subchronic administration of ketamine in subanaesthetic doses somewhat paradoxically elevates neuronal activity and neurotransmitter release in various brain regions. This effect has been recorded in both primates and rodents using several methods. For example, in humans, subanaesthetic infusion of ketamine during PET scan increased regional glucose metabolic rate in most brain regions, including the anterior and posterior cingulate cortex, caudate, and thalamus. No regions with decreased glucose metabolism were found [@langsjoEffectsSubanestheticKetamine2004]. Similarly, studies of brain haemodynamic activity in rats, monkeys, and humans all have shown quite consistent pattern of regional increases in the BOLD signal indicative of neuronal excitation [reviewed in @maltbieKetaminePharmacologicalImaging2017]. In all regions, peak BOLD response occurs 3–5 min after the start of ketamine infusion, which correlates well with the time-course of ketamine peak blood levels [@deakinGlutamateNeuralBasis2008]. In monkeys, acute ketamine injection increases both spontaneous and task-related firing of layer V pyramidal cells in the prefrontal cortex [@wangNMDAReceptorsSubserve2013]. There is experimental evidence that while the activity of the pyramidal cells in increased after ketamine infusion, the inhibition of NMDA receptors mainly decreases the activity of putative GABA interneurons [@homayounNMDAReceptorHypofunction2007]. Therefore at least some of the elevated cerebral excitation stems from the disinhibition of projection neurons that were released from GABA-ergic inhibition.

The effects of ketamine on neurotransmitter release are reviewed in [@napolitanoNeurometabolicProfilingKetamine2016]. Subchronic ketamine administration in rats releases glutamate in the frontal lobe [@chatterjeeNeurochemicalMolecularCharacterization2012; @kimVivoExVivo2011]. If ketamine is administered acutely, the release of glutamate in the prefrontal cortex is dose-dependent [@moghaddamActivationGlutamatergicNeurotransmission1997]. Similarly, acute and subchronic PCP administration increased glutamate release and reduced levels of GABA neuronal markers parvalbumin and GAD67 in the prefrontal cortex [@amitaiRepeatedPhencyclidineAdministration2012]. A number of other neurotransmitters have also been studied. Increased cerebral serotonin turnover after were observed after acute and subchronic ketamine administration [@chatterjeeNeurochemicalMolecularCharacterization2012; @vargiuPossibleRoleBrain1978]. Acute dizocilpine administration increased dopamine and serotonin turnover in the frontal cortex, piriform cortex, hippocampus, and striatum [@hiramatsuComparisonBehavioralBiochemical1989; @loscherNmethylDaspartateReceptorAntagonist1991]. Similar results have also been obtained with PCP [@hondoEffectPhencyclidineDopamine1994; @nabeshimaSerotonergicInvolvementPhencyclidineinduced1984]. 
Chatterjee and colleagues [-@chatterjeeNeurochemicalMolecularCharacterization2012] have performed a comprehensive research on acute and subchronic (10 days of treatment) effects of ketamine in mice. They found higher dopamine and acetylcholine turnover, but lower glycine levels in cortex, striatum, and hippocampus[@chatterjeeNeurochemicalMolecularCharacterization2012; see also @kikuchiEffectsKetaminePentobarbitone1997]. Gene expression was also affected. D~1~ and D~2~ dopamine, as well as 5-HT~1A~ receptor expressions in the cortex was significantly elevated. The gene-expression of two NMDAR subunits NR1 and NR2b was likewise higher in the cortex after subchronic ketamine administration [@chatterjeeNeurochemicalMolecularCharacterization2012]. Previously, the same group found that behavioural effects of subchronic ketamine regimen in mice persist for at least 10 days after the withdrawal of ketamine administration [@chatterjeeEffectChronicAcute2011]. Besides NMDAR, ketamine binds to dopamine D~2~ receptors in high affinity state, serotonin 5-HT~2A~, sigma 1, and opioid receptors [@frohlichReviewingKetamineModel2014; @kapurNMDAReceptorAntagonists2002]. The results we have reviewed show that introduction of ketamine rapidly reconfigures and elevates neuronal activity via several molecular target pathways. Finally, in some studies the metabolic activity was also measured from histological preparations similar to the current study. A quantitative autoradiographic ^14^C-2-deoxyglucose (2-DG) technique is used to measure the glucose oxidation just minutes after the acute dose of NMDAR antaganist. Pronounced increases in local cerebral
glucose utilisation were observed after acute dizocilpine dose in the olfactory areas and in a number of brain areas in the limbic system (including dentate gyrus, posterior cingulate cortex, and mammillary body) [@kurumajiEffectsMK801Local1989]. Similar results were found for ketamine (35 mg/kg). Increased 2-DG uptake was recorded in cingulate, retrosplenial, and piriform cortices, dentate gyrus, various thalamic nuclei, although there was no change in medial geniculate [@duncanMetabolicMappingRat1998]. Of note is that while in pharmacological studies PCP was quite similar to ketamine, on histological sections the reported results are opposite. Chronic and subchronic treatments of PCP were associated with reduced glucose metabolism in some prefrontal areas in the cortex and auditory centres of the brain [@cochranInductionMetabolicHypofunction2003]. Reduced brain metabolism after the chronic PCP regimen was also identified in COX activity, but mainly in basal ganglia and septum [@princeNormalizationCytochromecOxidase1997].

Differential correlation analysis is complimentary to the comparison of the mean COX levels. It accounts for the shape of the distribution of the results and allows to compare pairwise regional interactions between conditions. Four layers of the superior colliculi and one subdivision of the periaqueductal gray matter showed reduced positive correlations over all brain regions after subchronic ketamine treatment. Several layers of the superior colliculi were also prominently different between two rat groups in regional pairwise correlations of their COX levels. Their correlations with cortical and hippocampal brain regions were strongly positive in the control condition and tended to become negative after subchronic ketamine administration. Dorsomedial periaqueductal gray, hippocampal CA3, and prerubral field were three other brain regions with 3-4 pairwise significant reductions in positive correlations in the latter condition. 

The opposite pattern of changes in the pairwise correlation coefficients between two conditions was also observed, but in the fewer brain regions. None of these brain regions showed increased positive pairwise correlations after ketamine administration with more than 2 other brain regions. There was also little overlap between brain regions that exhibited more positive and more negative pairwise correlations in the ketamine group rats. The reduction of the positive pairwise correlations after subchronic ketamine administration was a more common pattern and various layers of the superior colliculi were prominent nodes in this process.

Extensive metabolic and extracellular changes in cortex, hippocampus, and basal ganglia after the administration of ketamine are already well documented [eg @chatterjeeNeurochemicalMolecularCharacterization2012; @duncanMetabolicMappingRat1998]. In the current study a significant role of the subcortical sensory regions have also emerged.  
Lateral posterior thalamic nucleus and superficial layers of superior colliculi (SuG(-5.8), Op(-5.8)) receive direct retinal projections and are important for visually guided behaviours and multisensory integration [@allenVisualInputMouse2016; @dragerTopographyVisualSomatosensory1976]. Some neurons in the deeper layers of the superior colliculus are also responsive to auditory stimulation. In cats, the acute infusion of ketamine facilitated spontaneous activity and auditory responses of neurons in the intermediate layers of the superior colliculus [@populinAnestheticsChangeExcitation2005]. Ventral part of the medial geniculate is the primary auditory relay nucleus of the thalamus that projects to the primary auditory cortex [@ryugoDifferentialTelencephalicProjections1974]. COX activity in MGV(−5.8) was elevated in absolute levels. Its correlations with Dp(-5.8) and DRD(−8.0) has also become significantly positive. It is noteworthy that acute doses of ketamine produce visual and auditory deficits in healthy humans and rodents [eg @hillhouseEffectsNoncompetitiveNmethylDaspartate2015; @umbrichtKetamineinducedDeficitsAuditory2000]. Prerubral field belongs to the H fields of Forel: a meeting place of several fibre bundles forming cortico-striato-thalamo-cortical loops where the convergence is observed of sensorimotor, associative, and limbic pathways [@neudorferNeuroanatomicalBackgroundFunctional2018]. In rat, prerubral field receives axonal projections from whisker-sensitive region of the spinal trigeminal nucleus [@veinanteThalamicProjectionsWhiskersensitive2000]. Finally, the olfactory circuitry was also represented via an increased COX levels in the piriform cortex [@alkoborssyModulationOlfactorydrivenBehavior2019]. 

The limitation of our study was relatively low number of the animals, therefore some significant effects of ketamine administration are likely to be masked by the presence of 1 or 2 outliers or missing observations. The activity of several sensory neurocircuits, especially in the superior colliculi, was affected by ketamine adminisration.

Histology remains the best way to measure neuronal activity in smaller subcortical brain regions. Our study is the first to evaluate the effect of ketamine administration on regional brain metabolic activity by means of histochemistry.
# Conclusion

# Abbreviations
## Brain regions
Brain regions are abbreviated according to the reference brain atlas [@paxinosRatBrainStereotaxic2007]. For each brain region the distance to bregma in millimetres according to the reference brain secion is provided in parentheses.
APT (-4.8), Anterior pretectal nucleus; AuD (-4.8), Auditory cortex, secondary, dorsal;
AuV (-4.8), Auditory cortex, secondary, ventral; B (-1.3), Nucleus basalis; CA1 (-4.8), Hippocampal CA1(Cornu Ammonis area 1); CA2 (-4.8), Hippocampal CA2; CA3 (-4.8), Hippocampal CA3; Cg1,2 (-0.3), Cingulate cortex, areas 1 & 2; CPu (-1.3), Caudate putamen; CPuDM (+0.2), Caudate putamen, dorsomedial; CPuV (+0.2), Caudate putamen, ventral; DG (-3.8), Hippocampal dentate gyrus; DG (-4.8), Hippocampal dentate gyrus; DLG (-4.8), Dorsal lateral geniculate; DMPAG (-5.8), Dorsomedial periaqueductal gray; DP (+2.2), Dorsal peduncular cortex; Dp (-5.8), Superior colliculi, deep gray/white layer; DRD (-8.0), Dorsal raphe, dorsal; Ect (-4.8), Ectorhinal cortex; GP (-1.3), Globus pallidus; InG (-5.8), Superior colliculi, intermediate gray layer; InWh (-5.8), Superior colliculi, intermediate white layer; LC (-10.04), Locus coeruleus; LP (-4.8), Lateral posterior nucleus; LPMR (-3.8), Lateral posterior thalamus, mediorostral; LSD (-0.3), Lateral septal nucleus, dorsal; M2 (-0.3), Motor cortex, secondary; MCPO (-0.3), Magnocellular preoptic nucleus; MGV (-5.8), Medial geniculate, ventral; MM (-4.8), Medial mamillary, medial; Op (-5.8), Superior colliculi, optic nerve layer; Pir (-1.3), Piriform cortex; PR (-4.8), Prerubral field; PtA (-4.8), Parietal association cortex; RSA (-5.8), Retrosplenial cortex, agranular; RSG (-4.8), Retrosplenial cortex, granular; RSGa (-5.8), Retrosplenial cortex, granular; RSGb (-5.8), Retrosplenial cortex, granular; RTg (-8.0), Reticulotegmental nucleus; Shi (+0.2), Septohippocampal nucleus; SPTg (-8.0), Subpeduncular tegmental nucleus; SubD (-5.8), Subiculum, dorsal; SuG (-5.8), Superior colliculi, superficial gray layer; V1M (-8.0), Primary visual cortex, monocular; V2 (-4.8), Visual cortex, secondary. 


## Other
2-DG, ^14^C-2-deoxyglucose; 5-HIAA, 5-hydroxyindoleacetic acid; AMPA, $\alpha$-amino-3-hydroxy-5-methylisoxazole-4-propionate; BOLD, blood oxygen level dependent;
COX, cytochrome c oxidase; DCA, differential correlation analysis; GABA, $\gamma$-aminobutyric acid; GAD67, glutamic acid decarboxylase-67; NMDAR, N-methyl-D-aspartate receptor; NRF-1, Nuclear respiratory factor 1; PCP, phencyclidine; PET, positron emission tomography; SE, standard error.

# Acknowledgements
This research has been supported by the Estonian Ministry of Education and Science project IUT20-40.

\pagebreak

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

\pagebreak


