---
title             : "Cytochrome c oxidase metabolic mapping in a chronic ketamine administration model"
shorttitle        : "cytochrome oxidase after chronic ketamine"
author:
  - name          : "Denis Matrov"
    affiliation   : "1,3"
  - name          : "Sophie Imbeault"
    affiliation   : "2"
  - name          : "Margus Kanarik"
    affiliation   : "3"
  - name          : "Marianna Shkolnaya"
    affiliation   : "2"
  - name          : "Patricia Schikorra"
    affiliation   : "2"
  - name          : "Ergo Miljan"
    affiliation   : "2"
  - name          : "Ruth Shimmo"
    affiliation   : "2"
  - name          : "Jaanus Harro"
    affiliation   : "3"  
    corresponding : yes
    address       : "Postal address"
    email         : "jaanus.harro@ut.ee"

affiliation:
  - id            : "1"
    institution   : "Department of Neuroscience, Graduate School of Medicine, Kyoto University, Kyoto, Japan"
  - id            : "2" 
    institution   : "Tallinn University Centre of Excellence in Neural and Behavioural Sciences, School of Natural Sciences and Health, Tallinn University, Tallinn, Estonia" 
  - id            : "3"
    institution   :  "Division of Neuropsychopharmacology, Department of Psychology, University of Tartu, Tartu, Estonia"
    
keywords          : ["rats", "cytochrome oxidase", "ketamine"]
wordcount         : "X"

bibliography      : ["R-packages.bib", "references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
csl               : "../templates/behavioural-brain-research.csl"
output            : 
  papaja::apa6_pdf:
    latex_engine: xelatex

header-includes:
- \usepackage{booktabs}
- \usepackage{makecell}
---

```{r setup, echo = FALSE, include = FALSE, cache=FALSE}
library("papaja")
library("ggplot2")
library("tidyverse")
library("scales")
library("gridExtra")
library("mice")
library("here")
library("kableExtra")

# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(
  cache = TRUE,
  cache.comments = FALSE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  echo = FALSE,
  strip.white = TRUE,
  comment = "#>",
  fig.path = "../figures/",
  results = "asis",
  tidy = "styler",
  dev = c('pdf', 'tiff'),
  dev.args = list(pdf = list(colormodel = 'cmyk'), 
                  tiff = list(compression = 'lzw')),
  dpi = 600,
  fig.width = 7,
  fig.height = 7,
  concordance = TRUE,
  global.par = TRUE
)

par(font.main = 1, cex.main = 1.05)
theme_set(theme_apa(base_size = 15) + theme(legend.position = "bottom"))
```



<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

# Introduction

Here is a citation

# Background


# Methods

```{r include-methods, child = "02-methods.Rmd"}
```



# Results

```{r include-results, child = "03-results.Rmd"}
```


# Discussion

# Conclusion

# Abbreviations
## Brain regions
Brain regions are abbreviated according to the reference brain atlas [@paxinosRatBrainStereotaxic2007]. For each brain region the distance to bregma in millimetres according to the reference brain secion is provided in parentheses.
APT(-4.8), Anterior pretectal nucleus; AuD(-4.8), Auditory cortex, secondary, dorsal;
AuV(-4.8), Auditory cortex, secondary, ventral; B(-1.3), Nucleus basalis; CA1(-4.8), Hippocampal CA1(Cornu Ammonis area 1); CA2(-4.8), Hippocampal CA2; CA3(-4.8), Hippocampal CA3; Cg1,2(-0.3), Cingulate cortex, areas 1 & 2; CPu(-1.3), Caudate putamen; CPuV(+0.2), Caudate putamen, ventral; DG(-3.8), Hippocampal dentate gyrus; DG(-4.8), Hippocampal dentate gyrus; DLG(-4.8), Dorsal lateral geniculate; DMPAG(-5.8), Dorsomedial periaqueductal gray; DP(+2.2), Dorsal peduncular cortex; Dp(-5.8), Superior colliculi, deep gray/white layer; DRD(-8.0), Dorsal raphe, dorsal; Ect(-4.8), Ectorhinal cortex; GP(-1.3), Globus pallidus; InG(-5.8), Superior colliculi, intermediate gray layer; InWh(-5.8), Superior colliculi, intermediate white layer; LC(-10.04), Locus coeruleus; LP(-4.8), Lateral posterior nucleus; LPMR(-3.8), Lateral posterior thalamus, mediorostral; LSD(-0.3), Lateral septal nucleus, dorsal; M2(-0.3), Motor cortex, secondary; MCPO (-0.3), Magnocellular preoptic nucleus; MGV(-5.8), Medial geniculate, ventral; MM(-4.8), Medial mamillary, medial; Op(-5.8), Superior colliculi, optic nerve layer; Pir(-1.3), Piriform cortex; PR(-4.8), Prerubral field; PtA(-4.8), Parietal association cortex; RSA(-5.8), Retrosplenial cortex, agranular (-5.8); RSG(-4.8), Retrosplenial cortex, granular (-4.8); RSGa(-5.8), Retrosplenial cortex, granular; RSGb(-5.8), Retrosplenial cortex, granular; RTg(-8.0), Reticulotegmental nucleus; SPTg(-8.0), Subpeduncular tegmental nucleus; SubD(-5.8), Subiculum, dorsal; SuG(-5.8), Superior colliculi, superficial gray layer; V1M(-8.0), Primary visual cortex, monocular; V2(-4.8), Visual cortex, secondary; 


## Other
SE, standard error; 

# Acknowledgements

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->

\pagebreak

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

\pagebreak

