---
title             : "Cytochrome c oxidase metabolic mapping in a chronic ketamine administration model"
shorttitle        : "cytochrome oxidase after chronic ketamine"
author:
  - name          : "Denis Matrov"
    affiliation   : "1,3"
  - name          : "Sophie Imbeault"
    affiliation   : "2"
  - name          : "Margus Kanarik"
    affiliation   : "3"
  - name          : "Marianna Shkolnaya"
    affiliation   : "2"
  - name          : "Patricia Schikorra"
    affiliation   : "2"
  - name          : "Ergo Miljan"
    affiliation   : "2"
  - name          : "Ruth Shimmo"
    affiliation   : "2"
  - name          : "Jaanus Harro"
    affiliation   : "3"  
    corresponding : yes
    address       : "Postal address"
    email         : "jaanus.harro@ut.ee"

affiliation:
  - id            : "1"
    institution   : "Department of Neuroscience, Graduate School of Medicine, Kyoto University, Kyoto, Japan"
  - id            : "2" 
    institution   : "Tallinn University Centre of Excellence in Neural and Behavioural Sciences, School of Natural Sciences and Health, Tallinn University, Tallinn, Estonia" 
  - id            : "3"
    institution   :  "Division of Neuropsychopharmacology, Department of Psychology, University of Tartu, Tartu, Estonia"
    
keywords          : ["rats", "cytochrome oxidase", "ketamine"]
wordcount         : "X"

bibliography      : ["R-packages.bib", "references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
csl               : "../templates/behavioural-brain-research.csl"
output            : 
  papaja::apa6_pdf:
    latex_engine: xelatex

header-includes:
- \usepackage{booktabs}
- \usepackage{makecell}
---

```{r setup, echo = FALSE, include = FALSE, cache=FALSE}
library("papaja")
library("ggplot2")
library("tidyverse")
library("scales")
library("gridExtra")
library("mice")
library("here")
library("kableExtra")
library("ggraph")
library("grid")
library("igraph")

# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(
  cache = TRUE,
  cache.comments = FALSE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  echo = FALSE,
  strip.white = TRUE,
  comment = "#>",
  fig.path = "../figures/",
  results = "asis",
  tidy = "styler",
  dev = c('pdf', 'tiff'),
  dev.args = list(pdf = list(colormodel = 'cmyk'), 
                  tiff = list(compression = 'lzw')),
  dpi = 600,
  fig.width = 7,
  fig.height = 7,
  concordance = TRUE,
  global.par = TRUE
)

par(font.main = 1, cex.main = 1.05)
theme_set(theme_apa(base_size = 15) + theme(legend.position = "bottom"))
```



<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

# Introduction
The term "schizophrenia" was first used by Bleuler in reference to the separation of function between personality, thinking, memory, and perception of those afflicted with this disorder [@stotz-ingenlathEpistemologicalAspectsEugen2000]. Although prevalence is comparatively low (~0.7%), the direct and indirect costs are high [@chongGlobalEconomicBurden2016], making finding an effective treatment important to both patients and society. Schizophrenia symptoms are divided into three categories: positive (hallucinations, delusions), negative (anhedonia, blunted affect, asociality), and cognitive (impaired executive function and memory). The degree of heritability is high, up to 80% [@harrisonGenesSchizophreniaRecent2003], however the underlying genetics are complex [@harrisonSchizophreniaGenesGene2005]. Many of the genes identified in recent genome-wide association studies are implicated either directly or indirectly in glutamate signaling [@schizophreniaworkinggroupofthepsychiatricgenomicsconsortiumBiologicalInsights1082014, @stotz-ingenlathEpistemologicalAspectsEugen2000]. 

Glutamate is the most ubiquitous neurotransmitter in mammalian brain: about 60% of neurons contain glutamate, and virtually all neurons have some type of glutamate receptor. Early pharmacological observations of dissociative anesthetics ketamine and phencyclidine (PCP) have lead to the hypothesis that glutamate plays a causal role in schizophrenia. These drugs affect the brain by blocking a major type of ionotropic glutamate receptor called N-methyl-D-aspartate receptor (NMDAR). Modulation of the postsynaptic NMDARs is crucial for activity-dependent synaptic plasticity and memory. By analogy, a hypofunction of NMDAR-mediated glutamatergic signaling was implicated in the pathophysiology of schizophrenia[@coyleGlutamateSchizophreniaDopamine2006]. Evidence for this hypothesis partly relies on the observation that when normal adults are administered an NMDAR antagonist, such as ketamine, they develop negative, cognitive and positive symptoms like those seen in schizophrenia [@adlerComparisonKetamineinducedThought1999, @krystalSubanestheticEffectsNoncompetitive1994, @newcomerKetamineinducedNMDAReceptor1999] while administration of these compounds to someone with schizophrenia exacerbates their symptoms [@lahtiEffectsKetamineNormal2001]. Furthermore, stimulating NMDAR-mediated signaling using agonists of the glycine modulatory site has been effective in alleviating some of the symptoms of schizophrenia in clinical trials [@heresco-levyPlacebocontrolledTrialDcycloserine2002; @tsaiDserineAddedAntipsychotics1998].
Despite the early enthusiasm, the therapeutic efficacy of the first generation of antipsychotic drugs that primarily modulate glutamatergic neurotransmission was not great [@buchananCognitiveNegativeSymptoms2007].

Both acute and chronic administration of the NMDAR antagonists such as ketamine, PCP, or dizocilpine (MK-801) replicate aspects of schizophrenia in animal models across the three symptom modalities. A subchronic treatment of 7 to 10 days is sufficient to impair declarative and working memory, as well as executive functions in rodents [@badoEffectsLowdoseDserine2011; @liPersistingCognitiveDeficits2011; @neillAnimalModelsCognitive2010]. Negative symptoms are replicated by a reduction of non-aggressive behaviour in the social interaction test [@beckerKetamineinducedChangesRat2004]. Positive symptoms are the hardest to model in rodents. Increased locomotor activity in the open field test supports the diverse effects of NMDAR blockade where some types of behaviour are positively modulated [@gamaEffectsOmega3Dietary2012].


# Materials and Methods

```{r include-methods, child = "02-methods.Rmd"}
```


# Results

```{r include-results, child = "03-results.Rmd"}
```


# Discussion
The decreased
brain volume in schizophrenia is especially pronounced in cortical areas, as
well as in the hippocampus (Levitt et al., 2010; Tamminga, Stan, & Wagner,
2010).

# Conclusion

# Abbreviations
## Brain regions
Brain regions are abbreviated according to the reference brain atlas [@paxinosRatBrainStereotaxic2007]. For each brain region the distance to bregma in millimetres according to the reference brain secion is provided in parentheses.
APT(-4.8), Anterior pretectal nucleus; AuD(-4.8), Auditory cortex, secondary, dorsal;
AuV(-4.8), Auditory cortex, secondary, ventral; B(-1.3), Nucleus basalis; CA1(-4.8), Hippocampal CA1(Cornu Ammonis area 1); CA2(-4.8), Hippocampal CA2; CA3(-4.8), Hippocampal CA3; Cg1,2(-0.3), Cingulate cortex, areas 1 & 2; CPu(-1.3), Caudate putamen; CPuDM(+0.2), Caudate putamen, dorsomedial; CPuV(+0.2), Caudate putamen, ventral; DG(-3.8), Hippocampal dentate gyrus; DG(-4.8), Hippocampal dentate gyrus; DLG(-4.8), Dorsal lateral geniculate; DMPAG(-5.8), Dorsomedial periaqueductal gray; DP(+2.2), Dorsal peduncular cortex; Dp(-5.8), Superior colliculi, deep gray/white layer; DRD(-8.0), Dorsal raphe, dorsal; Ect(-4.8), Ectorhinal cortex; GP(-1.3), Globus pallidus; InG(-5.8), Superior colliculi, intermediate gray layer; InWh(-5.8), Superior colliculi, intermediate white layer; LC(-10.04), Locus coeruleus; LP(-4.8), Lateral posterior nucleus; LPMR(-3.8), Lateral posterior thalamus, mediorostral; LSD(-0.3), Lateral septal nucleus, dorsal; M2(-0.3), Motor cortex, secondary; MCPO(-0.3), Magnocellular preoptic nucleus; MGV(-5.8), Medial geniculate, ventral; MM(-4.8), Medial mamillary, medial; Op(-5.8), Superior colliculi, optic nerve layer; Pir(-1.3), Piriform cortex; PR(-4.8), Prerubral field; PtA(-4.8), Parietal association cortex; RSA(-5.8), Retrosplenial cortex, agranular; RSG(-4.8), Retrosplenial cortex, granular; RSGa(-5.8), Retrosplenial cortex, granular; RSGb(-5.8), Retrosplenial cortex, granular; RTg(-8.0), Reticulotegmental nucleus; Shi(+0.2), Septohippocampal nucleus; SPTg(-8.0), Subpeduncular tegmental nucleus; SubD(-5.8), Subiculum, dorsal; SuG(-5.8), Superior colliculi, superficial gray layer; V1M(-8.0), Primary visual cortex, monocular; V2(-4.8), Visual cortex, secondary. 


## Other
DCA, differential correlation analysis; NMDAR, N-methyl-D-aspartate receptor; SE, standard error; 

# Acknowledgements

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->

\pagebreak

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

\pagebreak

